A ``Dream'' Interface Design

### 2014-11-26 dream-design/dream-design.tex
replace button picture
more edits

Many of these mobile devices today utilize touchscreens instead of physical keyboards or buttons. This allows for a larger display without increasing the size of the device. Furthermore, touchscreen keyboards and buttons have the ability to change their layout based on user input and disappear when not needed. To compensate for the lack of tactile feedback provided by physical keys and buttons, touchscreen devices often include aural feedback in the form of audible clicks from a speaker. Haptic feedback in the form of device vibrations is often included as well.

Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=6in]{buttons.jpg}
\caption{A simple representation demonstrating the idea of different subsets of buttons rising up on the same surface at different times.}
\label{buttons}
\end{figure}

@@

\section{Rationale}
This layout is designed so that IBSVI so that they can engage their spatial cognition, perception and sensing resources while interacting with touch screens. The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.

Feedback. In design, it is important to show the effect of an action. Without feedback, one is always wondering whether anything has hap- pened. Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing. Without feed- back, we turn equipment off at improper times or restart unnecessarily, losing all our recent work. Or we repeat the command and end up having the operation done twice, often to our detriment. Feedback is critical (the design of everyday things).

In the good old days of the telephone, before the American tele- phone system was divided among competing companies, before tele- phones were fancy and had so many features, telephones were de- signed with much more care and concern for the user. Designers at the Bell Telephone Laboratories worried a lot about feedback. The push buttons were designed to give an appropriate feel—tactile feedback. When a button was pushed, a tone was fed back into the earpiece so the user could tell that the button had been properly pushed. When the phone call was being connected, clicks, tones, and other noises gave the user feedback about the progress of the call. And the speaker's voice was always fed back to the earpiece in a carefully controlled amount, because the auditory feedback (called "sidetone") helped the person regulate how loudly to talk. All this has changed. We now have tele- phones that are much more powerful and often cheaper than those that existed just a few years ago—more function for less money. To be fair, these new designs are pushing hard on the paradox of technology: added functionality generally comes along at the price of added com- plexity. But that does not justify backward progress. 
Why are the modern telephone systems so difficult to learn and to use? Basically, the problem is that the systems have more features and less feedback. Suppose all telephones had a small display screen, not unlike the ones on small, inexpensive calculators. The display could be used to present, upon the push of a button, a brief menu of all the features of the telephone, one by one. When the desired one was encountered, the user would push another button to indicate that it should be invoked. If further action was required, the display could tell the person what to do. The display could even be auditory, with speech instead of a visual display. Only two buttons need be added to the 
telephone: one to change the display, one to accept the option on display. Of course, the telephone would be slightly more expensive. The tradeoff is cost versus usability.

>>>>

Many mobile devices today utilize touchscreens instead of physical keyboards or buttons. This allows for a larger display without increasing the size of the device. Furthermore, touchscreen keyboards and buttons have the ability to change their layout based on user input and disappear when not needed. To compensate for the lack of tactile feedback provided by physical keys and buttons, touchscreen devices often include aural feedback in the form of audible clicks from a speaker. Haptic feedback in the form of device vibrations is often included as well.

However, the literature suggests that overall feedback is still insufficient and remains a major usability issue with touchscreen devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=6in]{buttons.png}
\caption{A simple representation demonstrating the idea of different subsets of buttons rising up on the same surface at different times.}
\label{buttons}
\end{figure}

@@

\section{Rationale}
The driving force behind the creation and development of this interface design is feedback. Feedback is considered a critical concept in the field of interaction design. Donald Norman, the author of \textit{The Design of Everyday Things}, repeatedly articulates the importance of showing the effect of an action \cite{Norman02}. Without feedback, the user is always wondering whether anything has happened. In my ``dream'' interface design, the tactile feedback from the physical buttons serve to help the user understand when an action has been performed. Furthermore, the additional aural feedback of the design also assists in communicating the current state of the device to the user.

Another interaction design concept that Norman discusses is how systems with more options and less feedback can be very difficult to use \cite{Norman02}. My design includes ``intelligent'' physical buttons so that the device provides the user with the minimum number of options any given time. Since the buttons will only appear in situations when they are needed, it narrows the options of the user. This ultimately decreases the cognitive load of the user as they have less buttons from which to choose.

This layout is designed so that IBSVI can engage their spatial cognition, perception and sensing resources while interacting with touch screens. The design will only enable the grid-like system on-command so to not distract or annoy those who do not benefit from it.

Feedback. In design, it is important to show the effect of an action. Without feedback, one is always wondering whether anything has happened. Maybe the button wasn't pushed hard enough; maybe the machine has stopped working; maybe it is doing the wrong thing. Without feedback, we turn equipment off at improper times or restart unnecessarily, losing all our recent work. Or we repeat the command and end up having the operation done twice, often to our detriment. Feedback is critical.
